TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-03-13 paper/draft.tex
add refs

Advances in instrumentation with the lauch of several spectroscopic surveys in the past decade have dramatically increased the sample of known M dwarfs, and our potential to study M dwarf abundances on both exoplanet and galactic scales at a large statistical level.
The largest of such catalogs include the \citealt{West:2011} sample of 70,841 sources from the Sloan Digital Sky Survey (SDSS DR7; \citealt{Abazajian:2009}), and the \citealt{Yi:2014} and \citealt{Guo:2015} catalogs of 93,619 from the Large Sky Area Multi-Object Fiber Spectroscopic Telescope (LAMOST DR1; \citealt{Zhao:2012}), which each have determined spectral subtype classifications, H$\alpha$ equivalent width, and a number of optical molecular band index measurements. The launch of the (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III mission \citep{Eisenstein:2011} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy.

Why is there a problem determining M dwarf metallicities (and especially for high res \apogee\ spectra)? Emphasize issues: (1) models dont accurately reproduce data; (2) empirical standards are relative not absolute, and they are sparesely sampled. 
Methods for determining metallicity of M dwarfs: (1) from atmospheric models of M dwarfs, (2) empirical methods: spectroscopic or photometric calibrations with FGK pairs.

>>>>

Advances in instrumentation with the lauch of several spectroscopic surveys in the past decade have dramatically increased the sample of known M dwarfs, and our potential to study M dwarf abundances on both exoplanet and galactic scales at a large statistical level.
The largest of such catalogs include the \citealt{West:2011} sample of 70,841 sources from the Sloan Digital Sky Survey (SDSS DR7; \citealt{Abazajian:2009}), and the \citealt{Yi:2014} and \citealt{Guo:2015} catalogs of 93,619 from the Large Sky Area Multi-Object Fiber Spectroscopic Telescope (LAMOST DR1; \citealt{Zhao:2012}), which each have determined spectral subtype classifications, H$\alpha$ equivalent width, and a number of optical molecular band index measurements. The launch of the (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III mission \citep{Eisenstein:2011} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy, yet pipeline measurements \citep{Perez:2016} of \teff\ and \feh\ have only been determined precise to 100\,K and 0.18\,dex down to $T_{\rm eff}\sim3550\,$K \citep{Schmidt:2016}.

Why is there a problem determining M dwarf metallicities (and especially for high res \apogee\ spectra)? Emphasize issues: (1) models dont accurately reproduce data; (2) empirical standards are relative not absolute, and they are sparesely sampled. 
Methods for determining metallicity of M dwarfs: (1) from atmospheric models of M dwarfs, (2) empirical methods: photometric calibrations (\citealt{Johnson:2009}; \citealt{Schlaufman:2010}) or spectroscopic calibrations (\citealt{Terrien:2012}; \citealt{Rojas-Ayala:2012}; \citealt{Mann:2013a}; \citealt{Newton:2014}) with FGK pairs. Our approach is the latter.
